### 2007-01-08 Documentation/part1.tex
*** empty log message ***

TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED. JK: PLEASE LET ME KNOW WHEN {\tt example-tour.c} IS FIXED. MAYBE WE SHOULD  ALSO UPDATE THE PARAMTER VALUES [WW].

@@

glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\ \hspace*{4cm} \&glb\_num\_of\_exps);

@@

 glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0], \\

@@

The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
for the $\chi^2$ calculation. 
A very useful function to include energy calibration errors is
\begin{function}
\GLBNS{ShiftEnergyScale}
{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
as obtained with \GLB{glbGetEminEmax}.
\end{function}
DOES THAT ALSO WORK WITH VARIABLE BINS? WHAT IS THE MEAING OF EMIN AND EMAX? [WW]
For details on energy calibration errors, see \Sec~\ref{sec:rules}.
%
In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}. Additionally, the
functions \GLB{GetEminEmax}, \GLB{GetEnergyWindow}, \GLB{GetEnergyWindowBins}, and \GLB{GetNumberOfBins} from \Sec~\ref{sec:aedl_names} are useful.  We recommend that you familiarize yourself with these functions at this point.

@@

Further functions useful in the context of user-defined systematics are:
\begin{function}
\GLBNS{SetSysErrorsList} 
{\tt int glbSetSysErrorsList(int exp, int rule, int on\_off, const double *sys\_list)}
changes the systematical errors defined in the AEDL file in experiment {\tt exp} and 
rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
\end{function}
\begin{function}
\GLBNS{SetSysErrorsListPtr} 
{\tt double* glbSetSysErrorsListPtr(int exp, int rule, int on\_off)}
returns a pointer to the list of systematical errors as defined in the AEDL file in experiment {\tt exp} and 
rule {\tt rule}. Choose
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
\end{function}
In addition, in some cases the systematics minimization may not converge very fast or not at all, because it end up in a local minimum. In this case, one can change the starting values of the minimizer, \ie, the
starting point from which the local minimizer rolls into the local minimum.
\begin{function}
\GLBNS{SetSysStartingValuesList} 
{\tt int glbSetSysStartingValuesList(int exp, int rule, int on\_off, const double *sys\_list)}
sets the starting values of the local systematics minimizer in experiment {\tt exp} and 
rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
Usually, zero is used for all starting values. 
\end{function}
IS THAT CORRECT (THAT ZERO IS USUALLY USED)? [WW]
\begin{function}
\GLBNS{GetSysStartingValuesListPtr} 
{\tt double *glbGetSysStartingValuesListPtr(int exp, int rule, int on\_off, const double *sys\_list)}
returns the starting values of the local systematics minimizer in experiment {\tt exp} and 
rule {\tt rule}. This list can be obtained for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}.
\end{function}
A useful trick is often to use the minimum from the last run as starting values for the next one
if the input parameters are only slighltly changed. One can obtain the minimum from the last run 
from the last call of the $\chi^2$ function (\cf, {\tt example5.c}).

@@

The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can include the full $n$-parameter correlation with external input. It also demonstrates how these two compare to each other. One can easily read off this example that 
there is a substantial impact of the
correlation with oscillation parameters other than $\deltacp$. 
Note that it uses the function
\GLB{ChiNP} for arbitrary projections from the next section for the minimization over $\deltacp$.
In addition, there is one interesting feature in guessing the
oscillation parameters in this example: In order to avoid falling
into the wrong minimum, the fit value of $\deltacp$ is guessed from \figu{projex} (left). This quite sophisticated ``guessing'' is typical
for neutrino factories because of the $(\deltacp, \theta_{13})$-degeneracy, whereas it is for superbeams often sufficient
to use the true values. A strong indication for a wrong guessing 
are discontinuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another. In such cases, the starting point of the minimizer has to be adjusted to help it find the true minimum.

Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm31}, and \GLB{ChiDm21}, which
can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.

@@

In some cases, it is necessary to obtain information from the \AEDL\
files. Very basic information is supplied with
\begin{function}
\GLBNS{GetEminEmax}
{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
{\tt experiment}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindow}
{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
{\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindowBins}
{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfBins}
{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfSimBins}
{\tt int glbGetNumberOfSimBins(int exp)} returns the number of sampling points for the experiment {\tt exp}.
\end{function}
SHOULD THIS NOT BE CALLED {\tt glbGetNumberOfSamplingPoints}??? [WW]

>>>>

TOUR ITSELF (+EXAMPLE FILE) TO BE UPDATED (VALUES MAY HAVE CHANGED DUE TO AEDL FILE DEFINITION
CHANGES ETC. MAYBE ONLY COMMENT ABOUT THAT [WW].

@@

glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\ \hspace*{4cm} \&glb\_num\_of\_exps);

@@

 glbInitExperiment("NFstandard.glb",\&glb\_experiment\_list[0], \\

@@

The $\chi^2$ function {\tt glb\_chi\_function} may use a number of helper functions for
for the $\chi^2$ calculation. 
A very useful function to include energy calibration errors is
\begin{function}
\GLBNS{ShiftEnergyScale}
{\tt glbShiftEnergyScale(double b, double *rates\_in, double *rates\_out,int bins, double emin, double emax)}
shifts the energy scale in {\tt rates\_in} by the amount {\tt b} and stores the result in 
{\tt rates\_out}. The parameters {\tt emin} and {\tt emax} denote the minimal and maximal energy
as obtained with \GLB{glbGetEminEmax}. Requires constant energy bin widths!
\end{function}
For details on energy calibration errors, see \Sec~\ref{sec:rules}.
%
In addition, in order to implement user-defined systematics, one usually needs low-level access to the event rate vectors using \GLB{GetRuleRatePtr}, \GLB{GetSignalFitRatePtr}, \GLB{GetBGFitRatePtr} and possibly other functions described in \Sec~\ref{sec:event_rates}. Additionally, the
functions \GLB{GetEminEmax}, \GLB{GetEnergyWindow}, \GLB{GetEnergyWindowBins}, and \GLB{GetNumberOfBins} from \Sec~\ref{sec:aedl_names} are useful.  We recommend that you familiarize yourself with these functions at this point.

@@

Further functions useful in the context of user-defined systematics are:
\begin{function}
\GLBNS{SetSysErrorsList} 
{\tt int glbSetSysErrorsList(int exp, int rule, int on\_off, const double *sys\_list)}
changes the systematical errors defined in the AEDL file in experiment {\tt exp} and 
rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
\end{function}
\begin{function}
\GLBNS{GetSysErrorsListPtr} 
{\tt double* glbGetSysErrorsListPtr(int exp, int rule, int on\_off)}
returns a pointer to the list of systematical errors as defined in the AEDL file in experiment {\tt exp} and 
rule {\tt rule}. Choose
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
\end{function}
In addition, in some cases the systematics minimization may not converge very fast or not at all, because it end up in a local minimum. In this case, one can change the starting values of the minimizer, \ie, the
starting point from which the local minimizer rolls into the local minimum.
\begin{function}
\GLBNS{SetSysStartingValuesList} 
{\tt int glbSetSysStartingValuesList(int exp, int rule, int on\_off, const double *sys\_list)}
sets the starting values of the local systematics minimizer in experiment {\tt exp} and 
rule {\tt rule} to the values in {\tt sys\_list}. This change can be perfomed for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}. 
Usually, zero is used for all starting values. 
\end{function}
\begin{function}
\GLBNS{GetSysStartingValuesListPtr} 
{\tt double *glbGetSysStartingValuesListPtr(int exp, int rule, int on\_off, const double *sys\_list)}
returns the starting values of the local systematics minimizer in experiment {\tt exp} and 
rule {\tt rule}. This list can be obtained for
systematics on or systematics off by using {\tt GLB\_ON} or {\tt GLB\_OFF} for {\tt on\_off}.
\end{function}
A useful trick is often to use the minimum from the last run as starting values for the next one
if the input parameters are only slighltly changed. One can obtain the minimum from the last run 
from the last call of the $\chi^2$ function (\cf, {\tt example5.c}).

@@

The example on page~\pageref{ex:corrproj} demonstrates how one can obtain \figu{projex} (right) with keeping all parameters but $\deltacp$ fixed, as well as how one can include the full $n$-parameter correlation with external input. It also demonstrates how these two compare to each other. One can easily read off this example that 
there is a substantial impact of the
correlation with oscillation parameters other than $\deltacp$. 
Note that it uses the function
\GLB{ChiNP} for arbitrary projections from the next section for the minimization over $\deltacp$.
In addition, there is one interesting feature in guessing the
oscillation parameters in this example: In order to avoid falling
into the wrong minimum, the fit value of $\deltacp$ is guessed from \figu{projex} (left). This quite sophisticated ``guessing'' is typical
for neutrino factories because of the $(\deltacp, \theta_{13})$-degeneracy, whereas it is for superbeams often sufficient
to use the true values. A strong indication for a wrong guessing 
are discontinuous jumps in the projected $\chi^2$-function, where the minimizer jumps from one minimum to another. In such cases, the starting point of the minimizer has to be adjusted to help it find the true minimum.
%
Other examples for projections onto a parameter axis while keeping exactly one parameter fixed are \GLB{ChiTheta23}, \GLB{ChiDm31}, and \GLB{ChiDm21}, which
can be found in \Tab~\ref{tab:stdfunctions} on page~\pageref{tab:stdfunctions}.

Since the number of different parameter vectors used by \GLOBES\ may be a bit confusing, we summarize the most common parameter sets used in the calculations:
\begin{description}
\item[The simulated/true values] are the values set in \GLB{SetRates}.
\item[The fit values] are the values in the first parameter {\tt in} of {\tt glbChiTheta13} (and the other minimization functions) which are {\em fixed} by definition (such as $\theta_{13}$ for {\tt glbChiTheta13}). 
\item[The starting point/educated guess for the minimizer] are the values in the first parameter {\tt in} of {\tt glbChiTheta13} (and the other minimization functions) which are {\em free} by definition (such as all but $\theta_{13}$ for {\tt glbChiTheta13}). 
\item[The minimization result] of the marginalization process are the values in the second parameter {\tt out} of {\tt glbChiTheta13} (and the other minimization functions) which are {\em free} by definition (such as all but $\theta_{13}$ for {\tt glbChiTheta13}). The other values in {\tt out} still correspond to the fit values. 
\item[The position of external priors,] \ie, the best-fit values of external input, is set by \GLB{SetCenralValues}.
\item[The magnitude of external errors,] \ie, the errors themselves of external input, is set by \GLB{SetInputErrors}.
\end{description}

@@

In some cases, it is necessary to obtain information from the \AEDL\
files. Very basic information is supplied with
\begin{function}
\GLBNS{GetEminEmax}
{\tt int glbGetEminEmax(int experiment, double *emin, double *emax)} returns the
energy range [{\tt emin},{\tt emax}] covered by the binning as defined in the AEDL file for the experiment
{\tt experiment}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindow}
{\tt int glbGetEnergyWindow(int exp, int rule, double *low, double *high)} returns the
energy window [{\tt low},{\tt high}] defined in the AEDL file for the experiment
{\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetEnergyWindowBins}
{\tt int glbGetEnergyWindowBins(int exp, int rule, int *lowbin, int *highbin)} returns the
energy window in terms of bin numbers [{\tt lowbin},{\tt highbin}] corresponding to the energy window defined in the AEDL file for the experiment {\tt exp} and rule {\tt rule}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfBins}
{\tt int glbGetNumberOfBins(int exp)} returns the number of bins for the experiment {\tt exp}.
\end{function}
\begin{function}
\GLBNS{GetNumberOfSamplingPoints}
{\tt int glbGetNumberOfSamplingPoints(int exp)} returns the number of sampling points for the experiment {\tt exp}.
\end{function}
